Effect of quality-adjusted production function inputs

### 2015-06-24 Articles/Presentation_ESEE_2015/ESEE_2015.Rnw
Added notes about response() not working. This is for CST models only.

  modelsCST <- list()
  # Output elasticities based on cost share theorem
  alphaCST <- 0.3 # Fraction of national accounts spent on capital
  betaCST <- 0.7 # Fraction of national accounts spent on labor
  for (country in countries){
    form <- keep_summands(formulas[[country]][["unadjusted"]], n=-3) # Remove energy
    yVar <- as.name(summands(form, n=1, right=FALSE))
    kVar <- as.name(summands(form, n=1))
    lVar <- as.name(summands(form, n=2))
    tVar <- as.name(summands(form, n=3))
    data <- historicalData[[country]]
    # Cobb-Douglas
    cdModCST <- lm(log(eval(yVar)) - alphaCST*log(eval(kVar)) - betaCST*log(eval(lVar)) ~ eval(tVar), data = data)
    attr(cdModCST, "naturalCoeffs") <- data.frame(lambda=coef(cdModCST)[[2]],
                                                  logscale=coef(cdModCST)[[1]], 
                                                  scale=exp(coef(cdModCST)[[1]]),
                                                  alpha_1=alphaCST,
                                                  alpha_2=betaCST, 
                                                  alpha_3=0)
    cdModCST$formula <- form
    sdata <- subset(data, select=all.vars(form))
    sdata <- data[complete.cases(sdata), unique(c(all.vars(form), names(data)))]
    cdModCST$data <- sdata
    modelsCST[[country]][["unadjusted"]][["noE"]][["CD"]] <- cdModCST
    # CES
    cesModCST <- plm(log(eval(yVar) / (alphaCST*eval(kVar)^(-rho_1) + betaCST*eval(lVar)^(-rho_1))^(-1/rho_1)) ~ eval(tVar),
                     data = historicalData[[country]],
                     params=c(rho_1=0.8))
    names(cesModCST$coefficients) <- c("logscale", "lambda", "rho_1")
    cesModCST$coefficients <- c(coef(cesModCST), c(delta_1=alphaCST, delta=1))
    attr(cesModCST, "naturalCoeffs") <- EconModels:::makeNatCoef(cesModCST)
    cesModCST$formula <- form
    cesModCST$data <- sdata
    modelsCST[[country]][["unadjusted"]][["noE"]][["CES"]][["kl"]] <- cesModCST
  }

  # Save cost-share theorem models to disk.
  saveRDS(modelsCST, file=fileNameCST)

@@

<<Fitted_and_Residuals>>=
fittedAndResidData <- function(model, id){
  df <- data.frame(select(model$data, Year)) # getData doesn't work for lm models that are created for CST=TRUE.
  # By including [c(1:length(df$Year))] on the end, we 
  # add NAs (as appropriate) to the end of these columns
  # if data do not exist.
  df$fitted <- yhat(model)[c(1:length(df$Year))]
  df$Residuals <- resid(model)[c(1:length(df$Year))]
  # Now add the response, using the original variable name.
  df[[summands(model$formula, right = FALSE)]] <- response(model)[c(1:length(df$Year))]
  # Now add meta information
  id_list <- strsplit(id, split = "[.]")[[1]]
  df$Country <- id_list[[1]]
  df$flavor <- if (id_list[[2]] == "unadjusted") "Unadjusted" else "Quality-adjusted"
  df$energy <- if (id_list[[3]] == "noE") "Without energy" else "With energy"
  df$model <- id_list[[4]]
  df$nest <- if (length(id_list) > 4) id_list[[5]] else NA
  return(df)  
}

temp_data_CST <- leaf_apply(modelsCST, class=c("plm", "lm"), f = function(model, id){
  return(fittedAndResidData(model, id))
}) %>% do.call(rbind, .)
temp_data_CST$cst <- TRUE

temp_data_noCST <- leaf_apply(models, class=c("CDEmodel", "cesModel", "LINEXmodel"), f = function(model, id){
  return(fittedAndResidData(model, id))  
}) %>% do.call(rbind, .)
temp_data_noCST$cst <- FALSE

@@

<<Fitted and Residual Graph Functions>>=
fittedGraph <- function(data){
  ggplot() +
    geom_line(data=filter(data, energy=="Without energy"), mapping=aes(x=Year, y=iGDP)) +
    geom_line(data=data, mapping=aes(x=Year, y=fitted, linetype=energy)) +
    scale_linetype_manual(values=c(2,3), name="") +
    facet_grid(Country ~ flavor) +
    scale_x_continuous(breaks=c(1960, 1980, 2000)) +
    scale_y_continuous(breaks=c(1, 2, 4, 6, 8, 10, 12)) +
    xlab("") +
    ylab("Indexed GDP (1960=1)") +
    xy_theme()
}
@

Both historical GDP and fitted GDP are shown in 
Figures~\ref{fig:Graphs CD-fitted}--\ref{fig:Graphs CESekl-fitted}.
%
<<Graphs CD-fitted, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Cobb-Douglas models. Solid line is historical GDP.">>=
fittedGraph(filter(fitted_and_resid_data, model=="CD" & cst==FALSE))
@
%
<<Graphs CESkle-fitted, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="CES models with ($kl$)$e$ nesting. Solid line is historical GDP.">>=
fittedGraph(filter(fitted_and_resid_data, model=="CES" & cst==FALSE & (nest=="kle" | nest=="kl")))
@
%
<<Graphs CESlek-fitted, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="CES models with ($le$)$k$ nesting. Solid line is historical GDP.">>=
fittedGraph(filter(fitted_and_resid_data, model=="CES" & cst==FALSE & (nest=="lek" | nest=="kl")))
@
%
<<Graphs CESekl-fitted, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="CES models with ($ek$)$l$ nesting. Solid line is historical GDP.">>=
fittedGraph(filter(fitted_and_resid_data, model=="CES" & cst==FALSE & (nest=="ekl" | nest=="kl")))
@

>>>>

  modelsCST <- list()
  # Output elasticities based on cost share theorem
  alphaCST <- 0.3 # Fraction of national accounts spent on capital
  betaCST <- 0.7 # Fraction of national accounts spent on labor
  for (country in countries){
    form <- keep_summands(formulas[[country]][["unadjusted"]], n=-3) # Remove energy
    yVar <- as.name(summands(form, n=1, right=FALSE))
    kVar <- as.name(summands(form, n=1))
    lVar <- as.name(summands(form, n=2))
    tVar <- as.name(summands(form, n=3))
    data <- historicalData[[country]]
    # Cobb-Douglas
    cdModCST <- lm(log(eval(yVar)) - alphaCST*log(eval(kVar)) - betaCST*log(eval(lVar)) ~ eval(tVar), data = data)
    attr(cdModCST, "naturalCoeffs") <- data.frame(lambda=coef(cdModCST)[[2]],
                                                  logscale=coef(cdModCST)[[1]], 
                                                  scale=exp(coef(cdModCST)[[1]]),
                                                  alpha_1=alphaCST,
                                                  alpha_2=betaCST, 
                                                  alpha_3=0)
    cdModCST$formula <- form
    sdata <- subset(data, select=all.vars(form))
    sdata <- data[complete.cases(sdata), unique(c(all.vars(form), names(data)))]
    cdModCST$data <- sdata
    cdModCST$response <- sdata$iGDP 
    modelsCST[[country]][["unadjusted"]][["noE"]][["CD"]] <- cdModCST
    # CES
    cesModCST <- plm(log(eval(yVar) / (alphaCST*eval(kVar)^(-rho_1) + betaCST*eval(lVar)^(-rho_1))^(-1/rho_1)) ~ eval(tVar),
                     data = historicalData[[country]],
                     params=c(rho_1=0.8))
    names(cesModCST$coefficients) <- c("logscale", "lambda", "rho_1")
    cesModCST$coefficients <- c(coef(cesModCST), c(delta_1=alphaCST, delta=1))
    attr(cesModCST, "naturalCoeffs") <- EconModels:::makeNatCoef(cesModCST)
    cesModCST$formula <- form
    cesModCST$data <- sdata
    modelsCST[[country]][["unadjusted"]][["noE"]][["CES"]][["kl"]] <- cesModCST
  }
  # Save cost-share theorem models to disk.
  saveRDS(modelsCST, file=fileNameCST)

@@

<<Fitted_and_Residuals>>=
fittedAndResidData <- function(model, id){
  # ******************************************
  # getData doesn't work for lm models that are created for CST=TRUE.
  # After we implement fixed values for alpha_1 and alpha_2 in our fitting functions, 
  # getData(model) will probably work.
  # At that time, change model$data to getData(model)
  df <- data.frame(select(model$data, Year)) 
  # By including [c(1:length(df$Year))] on the end, we 
  # add NAs (as appropriate) to the end of these columns
  # if data do not exist.
  df$fitted <- yhat(model)[c(1:length(df$Year))]
  df$Residuals <- resid(model)[c(1:length(df$Year))]
  # Now add the response, using the original variable name.
  df[[summands(model$formula, right = FALSE)]] <- response(model)[c(1:length(df$Year))]
  # Now add meta information
  id_list <- strsplit(id, split = "[.]")[[1]]
  df$Country <- id_list[[1]]
  df$flavor <- if (id_list[[2]] == "unadjusted") "Unadjusted" else "Quality-adjusted"
  df$energy <- if (id_list[[3]] == "noE") "Without energy" else "With energy"
  df$model <- id_list[[4]]
  df$nest <- if (length(id_list) > 4) id_list[[5]] else NA
  return(df)  
}
temp_data_noCST <- leaf_apply(models, class=c("CDEmodel", "cesModel", "LINEXmodel"), f = function(model, id){
  return(fittedAndResidData(model, id))  
}) %>% do.call(rbind, .)
temp_data_noCST$cst <- FALSE

temp_data_CST <- leaf_apply(modelsCST, class=c("plm", "lm"), f = function(model, id){
  # ****************************************************
  # Not all of the results from this code are not correct at this time.
  # The residuals are OK.
  # But, the yhat() call in fittedAndResidData() will not return the correct information from the CST models.
  # This is because modelsCST was not created with our typical functions.
  # To correct this problem, we need to add arguments to cdModel and cesModel such that 
  # the models produced by those functions will be in the correct form for yhat() function.
  return(fittedAndResidData(model, id))
}) %>% do.call(rbind, .)
temp_data_CST$cst <- TRUE

@@

<<Fitted and Residual Graph Functions>>=
fittedGraphCST <- function(data){
  ggplot() +
    geom_line(data=filter(data, cst==TRUE), mapping=aes(x=Year, y=iGDP)) +
    geom_line(data=data, mapping=aes(x=Year, y=fitted, linetype=cst)) +
    scale_linetype_manual(values=c(2,3), name="") +
    facet_grid(Country ~ flavor) + 
    xlab("") +
    ylab("Indexed GDP (1960=1)") +
    xy_theme()
}

fittedGraphNoCST <- function(data){
  ggplot() +
    geom_line(data=filter(data, energy=="Without energy"), mapping=aes(x=Year, y=iGDP)) +
    geom_line(data=data, mapping=aes(x=Year, y=fitted, linetype=energy)) +
    scale_linetype_manual(values=c(2,3), name="") +
    facet_grid(Country ~ flavor) +
    scale_x_continuous(breaks=c(1960, 1980, 2000)) +
    scale_y_continuous(breaks=c(1, 2, 4, 6, 8, 10, 12)) +
    xlab("") +
    ylab("Indexed GDP (1960=1)") +
    xy_theme()
}
@

Both historical GDP and fitted GDP are shown in 
Figures~\ref{fig:Graphs CD-fitted}--\ref{fig:Graphs CESekl-fitted}.
%
<<Graphs CD-fitted, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Cobb-Douglas models that reject the cost-share theorem. Solid line is historical GDP.">>=
fittedGraphNoCST(filter(fitted_and_resid_data, model=="CD" & cst==FALSE))
@
%
<<Graphs CESkle-fitted, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="CES models with ($kl$)$e$ nesting. Solid line is historical GDP.">>=
fittedGraphNoCST(filter(fitted_and_resid_data, model=="CES" & cst==FALSE & (nest=="kle" | nest=="kl")))
@
%
<<Graphs CESlek-fitted, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="CES models with ($le$)$k$ nesting. Solid line is historical GDP.">>=
fittedGraphNoCST(filter(fitted_and_resid_data, model=="CES" & cst==FALSE & (nest=="lek" | nest=="kl")))
@
%
<<Graphs CESekl-fitted, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="CES models with ($ek$)$l$ nesting. Solid line is historical GDP.">>=
fittedGraphNoCST(filter(fitted_and_resid_data, model=="CES" & cst==FALSE & (nest=="ekl" | nest=="kl")))
@
